### 2020-03-06 abs.tex
almost there?

{\footnotesize We analyze a mechanistic model of a coronavirus-like pathogen
to assess the impact of asymptomatic cases on epidemic potential
-- as measured both by the basic reproduction
number and the fraction of new secondary cases attributable to asymptomatic
individuals. As we show, the impact of asymptomatic
cases depends on the interval over which asymptomatic cases
are infectious.  If the infectious duration of asymptomatic cases differs than that 
of symptomatic cases, then estimates of the basic reproductive ratio which do not explicitly account for asymptomatic cases may be systematically biased. Specifically, if asymptomatic cases have a shorter duration, 
${\cal{R}}_0$ will be over-estimated, and if they have a longer duration, ${\cal{R}}_0$ will be under-estimated.
%If infectiousness of asymptomatic cases is relatively short/long in duration compared
%to symptomatic cases, then
%estimates of the basic reproduction number may be over/under-estimates
%compared to estimates only attributing secondary cases to symptomatic cases.
The analysis provides a rationale for assessing the duration of asymptomatic cases 
of COVID-19 in addition to their prevalence in the population.
}

>>>>

{\footnotesize We analyze a mechanistic model of a coronavirus-like pathogen
to assess the impact of asymptomatic cases on epidemic potential
-- as measured both by the basic reproduction
number and the fraction of new secondary cases attributable to asymptomatic
individuals. As we show, the impact of asymptomatic
cases depends on the generation intervals of asymptomatic cases.  
If the generation intervals of asymptomatic cases differs than that 
of symptomatic cases, then estimates of the basic reproductive ratio which do not explicitly account for asymptomatic cases may be systematically biased. Specifically, if asymptomatic cases have a shorter generation interval, 
${\cal{R}}_0$ will be over-estimated, and if they have a longer generation interval, ${\cal{R}}_0$ will be under-estimated.
%If infectiousness of asymptomatic cases is relatively short/long in duration compared
%to symptomatic cases, then
%estimates of the basic reproduction number may be over/under-estimates
%compared to estimates only attributing secondary cases to symptomatic cases.
The analysis provides a rationale for assessing the duration of asymptomatic cases 
of COVID-19 in addition to their prevalence in the population.
}

### 2020-03-06 short.body.tex
almost there?

Here, given $1/r=7$ days, $\bar G_s=6$ days, and $\kappa_s=0.5$, we evaluate the impacts of intrinsic proportion of asymptomatic transmission $z$ on the realized relevance of asymptomatic infection, given variation in relative mean generation interval $\bar G_a/\bar G_s$ and relative generation-interval dispersion $\kappa_a/\kappa_s$ of asymptomatic cases.
We vary $\bar G_a/\bar G_s$ and $\kappa_a/\kappa_s$ from 0.5 to 1.5, one at a time, and infer $q$ as well as ${\cal{R}}_0$ compatible with an estimated value of $r$ from case data given the generation interval distributions of asymptomatic and asymptomatic cases.
Across the ranges of parameters we explore, the intrinsic proportion of asymptomatic transmission $z$ is similar to the realized proportion $q$ (Figure~\ref{fig.relevance}A,C).
When the mean asymptomatic generation interval is longer than the mean symptomatic generation interval, $q$ decreases because short generation intervals (i.e., fast transmission) drive transmission events during the growth phase.
On the other hand, assuming a wider asymptomatic generation interval (higher $\kappa_a$) increases $q$ because the amount of short generation intervals (i.e., the proportion of faster transmission) increases.

%Figure~\ref{fig.relevance}C shows that
%when $\gamma_a<\gamma_s$, then the strength ${\cal{R}}_0$ increases
%with $q$. 
%This increase arises because asymptomatic infections that resolve
%slower increase the duration of
%overall generation intervals (see~\citep{park_2019practical}). In contrast,
%Figure~\ref{fig.relevance}D show that
%when $\gamma_a>\gamma_s$ then
%the strength ${\cal{R}}_0$ decreases with $q$.  
%This decrease arises because asymptomatic infections that resolve
%faster decrease the duration of
%overall generation intervals (again, see~\citep{park_2019practical}). In both 
%cases, the dependence of ${\cal{R}}_0$ on $q$ is
%compared to a baseline of 
%${\cal{R}}_0\approx 3.1$ in a model with no secondary
%case production from asymptomatic infections (see filled
%circle at bottom left of Figure~\ref{fig.relevance}C-D.
%In other words, when ${\cal{R}}_0$ is estimated without explicitly accounting for asymptomatic spread, it can be over- or under- estimated depending on the relative duration of infection between symptomatic and asymptomatic individuals.

@@

The present findings highlight the need to characterize
the duration of asymptomatic cases and their consequences
not only for contact tracing but for constraining
uncertainty in the strength of the ongoing COVID-19 outbreak~\citep{park_preprint}.
The present findings are also consistent with
a recent generalization linking speed, strength, and
generation intervals, such that
for a given observed speed
increases in the mean generation interval imply
larger reproduction number~\citep{park_2019practical}. Hence, reports of individuals
with mild, persistent cases 
suggest the possibility that the mean generation interval for COVID-19
could be longer than estimated from symptomatic cases alone - possibly
increasing ${\cal{R}}_0$ and driving a larger fraction of secondary cases
(see Figure~\ref{fig.relevance} top row).
However, if asymptomatic cases tend to be shorter, then
current estimates of ${\cal{R}}_0$ may be over-estimates of the
underlying strength (see Figure~\ref{fig.relevance} bottom row).
Finally, it is important to note we focused here on the exponential
phase, but that the asymptomatic relevance $q$ is time-dependent, and will
vary given the population state structure of infected individuals.  Such
variation in the ratio of asymptomatic to symptomatic cases
will also impact estimates of case fatality rates insofar
as many asymptomatic cases are not counted, particularly early in an outbreak.
Future work might also consider the ways in which asymptomatic individuals
modulate resurgent epidemics in a networked metapopulation~\citep{watts_pnas2005}.

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0.pdf}\\
\mbox{}\\
\includegraphics[width=0.4\textwidth]{figheatmap_kappa.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_kappa_R0.pdf}
\caption{Relevance of asymptomatic transmission given variation in
relative transmission and probability that a given, newly infected individual is asymptomatic.  (Top) Long duration asymptomatic cases increase ${\cal{R}}_0$ with increasing relevance.
(Bottom) Short duration asymptomatic cases decrease ${\cal{R}}_0$ with increasing relevance.
Technically, renewal equations provide a route to connect
observed increases in case counts, i.e., ${\cal{R}}_0 = \frac{1}{M(-r)}$
where $M(-r)=\int_0^{\infty}\mathrm{d} a e^{-ra}g(a)$.
For the present model,
$M(r)=qM_E(-r)M_{I_a}(-r)+(1-q)M_E(-r)M_{I_s}(-r)$
where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
$M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
$M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
and $q$ is the fraction of realized new secondary cases caused by
asymptomatic individuals.  In all of these heatmaps, $\gamma_e=1/4$,
$\gamma_s=1/10$, and $r=1/7$. Here, the average duration
in symptomatic and asymptomatic states are $T_s=1/\gamma_s$ and 
$T_a=1/\gamma_a$, respectively. Here the filled circle denotes ${\cal{R}}_0\approx 3.12$
in the event of $p=0$ and $\beta_a=0$, and the open symbols are for $\beta_a/\beta_s=0.5$ and 
$p=0.5$ and $p=0.9$ for reference.
%Renewal equations provide a route to connect
%observed increases in case counts, i.e.,
%\begin{equation}
%{\cal{R}}_0 = \frac{1}{M(-r)}
%\end{equation}
%where $r$ is the speed of the disease spread and 
%$M(r)$ is related to 
%$g(a)$ (the normalized age of transmission) such that 
%\begin{equation}
%M(r)=\int_0^{\infty}\mathrm{d} a e^{ra}g(a).
%\end{equation}
%For the present model, the moment generating function is
%\begin{equation}
%M(r)=qM_E(r)M_{I_a}(r)+(1-q)M_E(r)M_{I_s}(r)
%\end{equation}
%where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
%$M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
%$M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
%and $q$ is the fraction of realized new secondary cases caused by
%asymptomatic individuals. 
%The relevance $q$ is an unknown feature of
\label{fig.relevance}}
\end{center}
\end{figure*}

>>>>

Here, given $1/r=7$ days, $\bar G_s=6$ days, and $\kappa_s=0.5$, we evaluate the impacts of intrinsic proportion of asymptomatic transmission $z$ on the realized relevance of asymptomatic infection, given variation in relative mean generation interval $\bar G_a/\bar G_s$ and relative generation-interval dispersion $\kappa_a/\kappa_s$ of asymptomatic cases.
We vary $\bar G_a/\bar G_s$ and $\kappa_a/\kappa_s$ from 0.5 to 1.5, one at a time, and infer $q$ as well as ${\cal{R}}_0$ compatible with an estimated value of $r$ from case data given the generation interval distributions of asymptomatic and asymptomatic cases.
Across the ranges of parameters we explore, the intrinsic proportion of asymptomatic transmission $z$ is similar to the realized proportion $q$ (Figure~\ref{fig.relevance}A,C).
When the mean asymptomatic generation interval is longer than the mean symptomatic generation interval, $q$ decreases because short generation intervals (i.e., fast transmission) caused by symptomatic individuals drive the transmission during the growth phase.
On the other hand, assuming a wide generation-interval distribution (higher $\kappa_a$) for asymptomatic individuals increases $q$ because the amount of short generation intervals (i.e., the proportion of faster transmission by asymptomatic individuals) increases.

Figure~\ref{fig.relevance}C shows that when $\bar G_s<\bar G_a$, then the strength ${\cal{R}}_0$ increases with $z$.
This increase arises because asymptomatic infections that have long generation intervals increase the duration of overall generation intervals (see~\citep{park_2019practical})
Conversely, when $\bar G_a > \bar G_s$ then the strength ${\cal{R}}_0$ decreases with $z$ because short generation intervals of asymptomatic cases reduce the overall mean generation interval.
Likewise, the amount of dispersion in the generation intervals of asymptomatic cases $\kappa_a$ affect the amount of dispersion in the overall generation intervals; 
therefore, depending on whether $\kappa_a$ is larger or smaller than $\kappa_s$, increase in the proportion of asymptomatic transmission $z$ can either decrease or increase the strength ${\cal{R}}_0$.
In other words, when ${\cal{R}}_0$ is estimated without explicitly accounting for asymptomatic spread, it can be over- or under- estimated depending on the relative mean and dispersion in generation intervals of asymptomatic and symptomatic individuals.

@@

The present findings highlight the need to characterize
the course of infection of asymptomatic cases and their consequences
not only for contact tracing but for constraining
uncertainty in the strength of the ongoing COVID-19 outbreak~\citep{park_preprint}.
The present findings are also consistent with
a recent generalization linking speed, strength, and
generation intervals, such that
for a given observed speed
increases in the mean generation interval imply
larger reproduction number~\citep{park_2019practical}. Hence, reports of individuals
with mild, persistent cases 
suggest the possibility that the mean generation interval for COVID-19
could be longer than estimated from symptomatic cases alone - possibly
increasing ${\cal{R}}_0$ and driving a larger fraction of secondary cases
(see Figure~\ref{fig.relevance} top row).
However, if asymptomatic cases tend to have shorter infection, then
current estimates of ${\cal{R}}_0$ may be over-estimates of the
underlying strength (see Figure~\ref{fig.relevance} bottom row).
Finally, it is important to note we focused here on the exponential
phase, but that the asymptomatic relevance $q$ is time-dependent, and will
vary given the population state structure of infected individuals.  Such
variation in the ratio of asymptomatic to symptomatic cases
will also impact estimates of case fatality rates insofar
as many asymptomatic cases are not counted, particularly early in an outbreak.
Future work might also consider the ways in which asymptomatic individuals
modulate resurgent epidemics in a networked metapopulation~\citep{watts_pnas2005}.

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0.pdf}\\
\mbox{}\\
\includegraphics[width=0.4\textwidth]{figheatmap_kappa.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_kappa_R0.pdf}
\caption{Relevance of asymptomatic transmission given variation in
probability that a given, newly infected individual is asymptomatic. (Top) Long/short generation intervals of asymptomatic cases increase/decrease ${\cal{R}}_0$ with decreasing/increasing relevance.
(Bottom) Wide/narrow generation intervals of asymptomatic cases decrease/increase ${\cal{R}}_0$ with increasing/decreasing relevance.
% Technically, renewal equations provide a route to connect
% observed increases in case counts, i.e., ${\cal{R}}_0 = \frac{1}{M(-r)}$
% where $M(-r)=\int_0^{\infty}\mathrm{d} a e^{-ra}g(a)$.
% For the present model,
% $M(r)=qM_E(-r)M_{I_a}(-r)+(1-q)M_E(-r)M_{I_s}(-r)$
% where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
% $M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
% $M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
% and $q$ is the fraction of realized new secondary cases caused by
% asymptomatic individuals.  In all of these heatmaps, $\gamma_e=1/4$,
% $\gamma_s=1/10$, and $r=1/7$. Here, the average duration
% in symptomatic and asymptomatic states are $T_s=1/\gamma_s$ and 
% $T_a=1/\gamma_a$, respectively. Here the filled circle denotes ${\cal{R}}_0\approx 3.12$
% in the event of $p=0$ and $\beta_a=0$, and the open symbols are for $\beta_a/\beta_s=0.5$ and 
% $p=0.5$ and $p=0.9$ for reference.
%Renewal equations provide a route to connect
%observed increases in case counts, i.e.,
%\begin{equation}
%{\cal{R}}_0 = \frac{1}{M(-r)}
%\end{equation}
%where $r$ is the speed of the disease spread and 
%$M(r)$ is related to 
%$g(a)$ (the normalized age of transmission) such that 
%\begin{equation}
%M(r)=\int_0^{\infty}\mathrm{d} a e^{ra}g(a).
%\end{equation}
%For the present model, the moment generating function is
%\begin{equation}
%M(r)=qM_E(r)M_{I_a}(r)+(1-q)M_E(r)M_{I_s}(r)
%\end{equation}
%where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
%$M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
%$M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
%and $q$ is the fraction of realized new secondary cases caused by
%asymptomatic individuals. 
%The relevance $q$ is an unknown feature of
\label{fig.relevance}}
\end{center}
\end{figure*}
